\begin{thebibliography}{99}

\bibitem{Ab} R. Abraham, J.E. Marsden and T. Ratiu,
{\bfseries Manifolds, Tensor Analysis, and Applications},
Addison-Wesley Publishing Company, Inc., 1983.

\bibitem{Ad} R. A. Adams and J. Fournier,
{\bfseries Sobolev Spaces, 2$^{nd}$ Edition}, Academic Press, 2003.

\bibitem{Br} H. Brezis,
{\bfseries Analyse fonctionnelle: Theorie et applications}, Masson, 1983.

\bibitem{ChMa} A. J. Chorin and J. E. Marsden,
{\bfseries A Mathematical Introduction to Fluid Mechanics},
Springer-Verlag, Umiversitext, 1979.

\bibitem{Co} J. B. Conway,
{\bfseries Functions of One Complex Variable}, Springer-Verlag, 1978.

\bibitem{Daf} C. M. Dafermos, {\bfseries Hyperbolic Conservations Laws
in Continuum Physics, 4$^{th}$ Edition}, Springer-Verlag, 2016.

\bibitem{Flem} W. Fleming, {\bfseries Functions of Several Variables},
Springer-Verlag, 1977.

\bibitem{FoRA} G. D. Folland, {\bfseries Real Analysis}, John-Wiley,
1984.

\bibitem{FoAC} G. D. Folland, {\bfseries Advanced Calculus}, Printice
Hall, 2002.

\bibitem{FoPDE} G. D. Folland,
{\bfseries An Introduction to Partial Differential Equations, 2$^{nd}$ Edition},
Princeton Academic Press, 1995.

\bibitem{Ha} Jack K.\ Hale, {\bfseries Ordinary Differential equations},
Robert E.\ Krieger Publishing Company, 1969.

\bibitem{He} D. Henry,
{\bfseries Geometric Theory of Semilinear Parabolic Equations},
Lecture Notes in Mathematics, Springer-Verlag, 1974.

\bibitem{Ho} F. Hoppensteadt,
{\bfseries Mathematical Theories of Populations: Demographics,
Genetics and Epidemics}, SIAM, 1975.

\bibitem{Horm} L. Hörmander,
{\bfseries Linear Partial Differential Operators}, Springer-Verlag, 1963.

\bibitem{Iser} A. Iserles,
{\bfseries A First Course in the Numerical Analysis of Differential
Equations}, Cambridge University Press, 1996.

\bibitem{J} F. John,
{\bfseries Partial Differential Equations, $4^{th}$ edition},
Springer-Verlag, 1982.

\bibitem{Lax} P. D. Lax,
{\bfseries Hyperbolic Systems of Conservation Laws and the
Mathematical Theory of Shock Wave}, CBMS no. 11, SIAM, 1973.

\bibitem{Lax2} P. D. Lax,
{\bfseries The Formation and Decay of Shock Wave}, Amer.\ Math.\ Monthly, 79,
pp 227-241, 1972.

\bibitem{LeF} P. G. LeFloch,
{\bfseries Hyperbolic System of Conservation Laws: The Theory of
Classical and Nonclassical Shock Waves}, Springer Basel AG, 2002.

\bibitem{LiMa1} J. L. Lions and E. Magenes,
{\bfseries Non-Homogeneous Boundary Value Problems and Applications:
Vol.\ I}, Springer-Verlag, 1972.

\bibitem{LiMa2} J. L. Lions and E. Magenes,
{\bfseries Non-Homogeneous Boundary Value Problems and Applications:
Vol.\ II}, Springer-Verlag, 1972.

\bibitem{McO} R. C. McOwen,
{\bfseries Partial Differential Equations: Methods and Applications,
2$^{nd}$ Edition}, Prentice Hall, 2003.

\bibitem{Mu} J. R. Munkres,
{\bfseries Topology, A First Course}, Prentice Hall, 1975.

\bibitem{PinRub} Y. Pinchover and J. Rubinstein,
{\bfseries An Introduction to Partial differential Equations},
Cambridge University Press, 2005.

\bibitem{ProWei} M. Protter and H. Weinberger,
{\bfseries Maximum Principle in Differential Equations}, Prentice-Hall, 1967.

\bibitem{Rain} E. D. Rainville, {\bfseries Special Functions}, Chelsea
  Publishing Company, 1971.

\bibitem{ReeSim} M. Reed and B. Simon,
{\bfseries Methods of Modern Mathematical Physics, Vol.\ I: Functional
Analysis}, Academic Press, 1980.

\bibitem{RenRog} M. Renardy and R. C. Rogers,
{\bfseries An Introduction to Partial Differential Equations},
Springer-Verlag, 1993.

\bibitem{RuPMA} W. Rudin, {\bfseries Principles of Mathematical Analysis},
McGraw-Hill, 1976.

\bibitem{Ru} W. Rudin, {\bfseries Real and Complex Analysis},
McGraw-Hill, 1974.

\bibitem{RuFA} W. Rudin, {\bfseries Functional Analysis},
McGraw-Hill, 1973.

\bibitem{Sal} S. Salsa, {\bfseries Partial Differential Equations in
Action: From Modelling to Theory}, Springer-Verlag, 2008. 

\bibitem{Simm} G. F. Simmons, {\bfseries Differential Equations with
Applications and Historical Notes}, McGraw-Hill, 1972. 

\bibitem{Smi} G. D. Smith,
{\bfseries Numerical Solution of Partial Differential Equations,
2$^{nd}$ Edition}, Oxford University Press, 1978. 

\bibitem{Smo} J. Smoller,
{\bfseries Shock Waves and Reaction-Diffusion Equations, 2$^{nd}$ Edition},
Springer-Verlag, 1994.

\bibitem{Sob} S. L. Sobolev,
{\bfseries Partial Differential Equations of Mathematical Physics,
3$^{rd}$ Edition}, Pergamon Press, Addison-Wesley Publishing Company,
Inc, 1964.

\bibitem{Sp} M. Spivak,
{\bfseries A Comprehensive Introduction to Differential Geometry,
  Volume One, 3$^{rd}$ Edition}, Publish or Perish, Inc, 2005. 

\bibitem{Str} W. A. Strauss,
{\bfseries Partial Differential Equations: an Introduction, 2$^{nd}$ Edition},
John-Wiley \& Sons, 2008.

\bibitem{Tr} F. Treves,
{\bfseries Basic Linear Partial Differential Equations},
Academic Press, Inc., 1975.

\end{thebibliography}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "notes"
%%% End:
